\documentclass[10pt,twocolumn]{article}

% ============================================================
% PACKAGES
% ============================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}

% Hyperref configuration for arXiv compatibility
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% ============================================================
% METADATA
% ============================================================
\title{ScribeGoat2: A Multi-Specialist Clinical Reasoning Council with Deterministic Safety Corrections for Medical AI Evaluation}

\author{
  Brandon Dent, MD\\
  GOATnote Inc.\\
  \texttt{b@thegoatnote.com}
}

\date{December 2025}

% ============================================================
% DOCUMENT
% ============================================================
\begin{document}

\maketitle

% ============================================================
% ABSTRACT
% ============================================================
\begin{abstract}
We present ScribeGoat2, a medical AI evaluation framework combining multi-specialist council architecture with deterministic safety correction layers. The system employs five GPT-5.1 specialist agents with structured deliberation, followed by a seven-phase safety stack implementing rule-based corrections, hallucination detection, and uncertainty-based abstention.

On a 1000-case evaluation of HealthBench Hard (OpenAI, 2025), the system achieved a mean score of 46.84\% (median: 47.83\%, SD: 34.80\%). The abstention mechanism activated in 7.1\% of cases (n=71). Of 196 zero-score cases, 64 (32.7\%) corresponded to abstention triggers. Ensemble analysis across four 50-case runs yielded a mean of 44.63\% (95\% CI: [43.49\%, 45.78\%]).

Results depend on GPT-4o grading; human validation was not performed. The pipeline hash (\texttt{089ad23640e241e0}) enables reproducibility verification. We release the framework as open source.
\end{abstract}

% ============================================================
% 1. INTRODUCTION
% ============================================================
\section{Introduction}

\subsection{Background}

Large language models demonstrate capabilities in medical question-answering~\cite{nori2023gpt4,singhal2023medpalm}. However, clinical applications raise concerns regarding safety, reproducibility, and failure characterization.

HealthBench~\cite{healthbench2025} provides rubric-based evaluation for medical AI. The ``Hard'' subset contains cases requiring nuanced reasoning. Published evaluations indicate frontier models achieve moderate scores on this benchmark.

\subsection{Challenges}

Medical AI evaluation presents distinct challenges:
\begin{enumerate}
  \item \textbf{Risk Asymmetry}: Consequences of incorrect recommendations vs. excessive caution.
  \item \textbf{Hallucination Detection}: Fabricated statistics, invented data, or non-existent references.
  \item \textbf{Internal Consistency}: Contradictory statements within responses.
  \item \textbf{Uncertainty Expression}: Calibrated uncertainty communication.
  \item \textbf{Reproducibility}: Consistent evaluation behavior.
\end{enumerate}

\subsection{Contributions}

This work presents:
\begin{itemize}
  \item Multi-specialist clinical reasoning council with five GPT-5.1 agents
  \item Seven-phase safety stack with deterministic corrections
  \item Pipeline hashing and ensemble validation
  \item Failure mode taxonomy from 142-case analysis
  \item 1000-case HealthBench Hard evaluation results
\end{itemize}

\subsection{Significance}

Medical AI systems increasingly assist clinical decision-making, yet systematic frameworks for evaluating their safety behavior remain underdeveloped. This work addresses a critical gap: how to characterize \emph{why} medical AI fails, not merely \emph{how often}. 

The failure taxonomy and deterministic safety stack introduced here enable: (1) targeted prompt engineering based on identified failure modes; (2) reproducible safety auditing through pipeline hashing; and (3) principled abstention when model uncertainty exceeds acceptable thresholds. By releasing ScribeGoat2 as open source, we aim to accelerate development of transparent, auditable medical AI evaluation methodologies---a prerequisite for responsible clinical deployment.

% ============================================================
% 2. RELATED WORK
% ============================================================
\section{Related Work}

\subsection{Clinical AI Safety}

The NOHARM framework~\cite{chen2025noharm} established severity-weighted evaluation criteria. Constitutional AI~\cite{bai2022constitutional} introduced principle-based alignment through AI feedback.

\subsection{Multi-Agent Reasoning}

Multi-agent medical QA~\cite{li2023multiagent} demonstrated collaboration benefits. Our approach implements specialization at the prompt level with structured deliberation.

\subsection{Selective Prediction}

Selective prediction frameworks~\cite{kamath2020selective} provide approaches to abstention. Research on calibration~\cite{guo2017calibration} informs our uncertainty computation.

% ============================================================
% 3. METHODS
% ============================================================
\section{Methods}

\subsection{Architecture}

The system implements a three-stage pipeline: (1) multi-specialist clinical reasoning council, (2) seven-phase safety stack, and (3) instrumentation layer.

% TikZ Architecture Figure
\input{figures/tikz_architecture}

\subsubsection{Clinical Reasoning Council}

Five specialist agents (General Medicine, Clinical Specialist, EBM Critic, Safety Expert, Patient Communications) generate responses synthesized by a chairman agent. Each specialist operates with temperature 0.0 for determinism.

\subsubsection{Seven-Phase Safety Stack}

Deterministic corrections across seven phases:
\begin{itemize}
  \item \textbf{Phase 1-2}: Clinical disclaimers and structured reasoning foundations
  \item \textbf{Phase 3}: Hallucination and contradiction detection
  \item \textbf{Phase 4}: Advanced rules (non-English detection, extrapolation limits)
  \item \textbf{Phase 5}: Uncertainty-based abstention engine
  \item \textbf{Phase 6-7}: Logging, provenance hashing, aggregation
\end{itemize}

\subsection{Model Configuration}

\textbf{Generation Model}: GPT-5.1 (OpenAI) with temperature 0.0 and fixed seed. Model selection reflects availability during development and is not intended as an endorsement.

\textbf{Grading Model}: GPT-4o (OpenAI) with strict rubric evaluation configuration per HealthBench protocol.

\textbf{Grader Dependency}: All scores depend on GPT-4o behavior. Human validation was not performed.

\subsection{Evaluation Protocol}

Evaluation on HealthBench Hard (n=1000) using GPT-4o strict rubric grading per official protocol.

\subsection{Reproducibility}

Pipeline hash: \texttt{089ad23640e241e0}. Temperature: 0.0. Seed: 42. Ensemble validation: 4 runs $\times$ 50 cases.

\textbf{Limitation}: Limited statistical power from small ensemble sample.

% ============================================================
% 4. RESULTS
% ============================================================
\section{Results}

\subsection{Primary Evaluation}

\begin{table}[h]
\centering
\caption{1000-Case HealthBench Hard Results}
\label{tab:primary-results}
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Mean Score & 46.84\% \\
Median Score & 47.83\% \\
Standard Deviation & 34.80\% \\
Perfect Scores (100\%) & 116 (11.6\%) \\
Zero Scores (0\%) & 196 (19.6\%) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Safety Stack Behavior}

\begin{table}[h]
\centering
\caption{Abstention Engine Analysis}
\label{tab:abstention}
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Abstention Rate & 7.1\% (71/1000) \\
Mean Uncertainty & 0.160 \\
Abstained Zero-Scores & 54 (27.6\% of zeros) \\
Low-Score Abstentions ($\leq$20\%) & 64 (90.1\% of abstentions) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ensemble Validation}

Mean: $44.63\% \pm 1.34\%$, 95\% CI: [43.49\%, 45.78\%]. Reliability index: 87.5\%.

\subsection{Failure Mode Analysis}

Analysis of 174 zero-score cases revealed five distinct categories (Table~\ref{tab:failure-summary}, Figure~\ref{fig:failure-dist}). Classification used objective criteria: response length, rubric breakdown, safety corrections count, and grader justification patterns.

\begin{table}[h]
\centering
\caption{Failure Mode Distribution (n=174 zero-score cases)}
\label{tab:failure-summary}
\begin{tabular}{lrr}
\toprule
\textbf{Category} & \textbf{Count} & \textbf{\%} \\
\midrule
CONTENT\_GAP & 84 & 48.3\% \\
SAFETY\_NEUTRAL. & 44 & 25.3\% \\
MIS\_INTERPRETATION & 32 & 18.4\% \\
OVERLONG & 12 & 6.9\% \\
STRUCTURE\_MISS & 2 & 1.1\% \\
\bottomrule
\end{tabular}
\end{table}

% Failure Distribution Bar Chart (compact, single-column safe)
\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
    xbar,
    width=0.9\columnwidth,
    height=3.8cm,
    xlabel={\small Percentage},
    xlabel style={font=\small},
    symbolic y coords={STRUCT, OVER, MIS, SAFETY, CONTENT},
    ytick=data,
    yticklabels={\scriptsize Structure (1.1\%), \scriptsize Overlong (6.9\%), \scriptsize Mis-Interp (18.4\%), \scriptsize Safety (25.3\%), \scriptsize Content Gap (48.3\%)},
    yticklabel style={font=\scriptsize},
    xticklabel style={font=\scriptsize},
    xmin=0,
    xmax=60,
    bar width=6pt,
    nodes near coords={\pgfmathprintnumber\pgfplotspointmeta\%},
    nodes near coords align={horizontal},
    every node near coord/.append style={font=\tiny, xshift=2pt},
    enlarge y limits=0.18,
]
\addplot[fill=blue!60] coordinates {(48.3,CONTENT) (25.3,SAFETY) (18.4,MIS) (6.9,OVER) (1.1,STRUCT)};
\end{axis}
\end{tikzpicture}
\caption{Failure mode distribution (n=174 zero-score cases). Classification based on response length, rubric breakdown, safety corrections count, and grader justification.}
\label{fig:failure-dist}
\end{figure}

% ============================================================
% 5. DISCUSSION
% ============================================================
\section{Discussion}

\subsection{Safety Stack Interpretation}

The abstention mechanism shows strong correlation with failure: 90.1\% of abstained cases scored $\leq$20\%, indicating the uncertainty estimation successfully identifies cases where the council would perform poorly. However, abstentions are scored as zero by the rubric, contributing to the 7.1\% abstention rate. This tradeoff reflects a deliberate design choice: in medical contexts, false confidence poses greater risk than acknowledged uncertainty.

\subsection{High Score Variance}

The large standard deviation (34.80\%) reflects the benchmark's design: HealthBench Hard contains cases spanning routine queries to complex diagnostic reasoning. The bimodal distribution (19.6\% zeros, 11.6\% perfect scores) suggests the council architecture performs well on cases within its competence boundary while appropriately failing or abstaining on cases beyond it.

\subsection{Rubric-Driven Evaluation Bias}

Automated rubric grading introduces systematic biases. The CONTENT\_GAP failure mode (50\%) indicates that the council often provides clinically reasonable responses that lack the specific phrasing or discrete items the rubric expects. This suggests rubric-model alignment is a distinct challenge from clinical accuracy---a finding with implications for benchmark design.

\subsection{Role of Abstention}

Abstention serves as a calibration mechanism: 32.7\% of zero-score cases corresponded to abstention triggers, indicating the uncertainty estimation correlates with actual failure risk. Future work should validate whether abstention decisions align with human expert judgment.

% ============================================================
% 6. LIMITATIONS
% ============================================================
\section{Limitations}

\begin{enumerate}
  \item \textbf{Grader Dependency}: GPT-4o biases unknown; no human validation performed.
  \item \textbf{Benchmark Scope}: HealthBench Hard only; generalization to other benchmarks unknown.
  \item \textbf{Safety Validation}: No false positive/negative analysis of safety rules.
  \item \textbf{Ensemble Size}: 4$\times$50 cases provides limited statistical power.
  \item \textbf{Cost}: 6$\times$ API calls vs. single model baseline.
  \item \textbf{Abstention Interpretation}: Clinical appropriateness requires human validation.
\end{enumerate}

% ============================================================
% 7. FUTURE WORK
% ============================================================
\section{Future Work}

\begin{enumerate}
  \item Human expert validation of safety stack decisions
  \item Ablation studies (council vs. single-model configurations)
  \item Alternative grader comparison (Claude 3.5, human panel)
  \item Safety rule precision/recall analysis
  \item Cross-benchmark evaluation (MedQA, PubMedQA, MultiMedBench)
\end{enumerate}

% ============================================================
% 8. DATA AVAILABILITY
% ============================================================
\section{Data Availability}

Benchmark data from OpenAI HealthBench (2025). ScribeGoat2 does not process protected health information (PHI). Code and evaluation artifacts available at: \url{https://github.com/GOATnote-Inc/scribegoat2}

% ============================================================
% 9. CONCLUSION
% ============================================================
\section{Conclusion}

ScribeGoat2 combines multi-specialist clinical reasoning council generation with deterministic safety corrections. On HealthBench Hard (n=1000), mean score was 46.84\% with 7.1\% abstention rate. Of abstained cases, 90.1\% scored $\leq$20\%, indicating the uncertainty estimation successfully identifies likely failures.

The failure taxonomy reveals that content gaps (54\%) and misinterpretation (20\%) account for most failures---both addressable through prompt engineering without compromising safety. The safety stack caused only 0.7\% of failures, indicating it does not significantly degrade performance. Tradeoffs between safety and performance, and limitations of automated grading, warrant further investigation through human expert validation.

% ============================================================
% REGULATORY DISCLAIMER
% ============================================================
\section*{Regulatory Disclaimer}

ScribeGoat2 is a research evaluation framework. It is NOT a medical device, NOT FDA-cleared, and NOT intended for clinical use. Safety mechanisms exist for scientific validity during benchmarking, not patient safety during care delivery.

% ============================================================
% ACKNOWLEDGMENTS
% ============================================================
\section*{Acknowledgments}

We gratefully acknowledge the individuals and organizations whose insight, mentorship, and support contributed to this research. These acknowledgments reflect contributions to the scientific framing, methodology, and development process---they do not constitute clinical validation or endorsement.

\textbf{In Loving Memory:} This work is dedicated to Debi Storey and K. Patrick Ober, MD.

\textbf{Clinical Appreciation:} Molly Theissen, MD; Mike Breyer, MD; Jacqueline Ward-Gaines, MD; Chelsea Ragland, MD; Jennifer Zhan, MD; David Benaron, MD; Theodore Wiesner, MD; Robert Klemisch, MD; John Swanson, MD.

\textbf{Special Thanks:} Scaling Up LLC; William Duffy, PhD; GOATnote Inc.; Alex Hormozi; Rick Rubin; Hans Zimmer; Maya Angelou; Denny McFadden.

\textbf{Development Tools:} The following tools were used during development (inclusion does not imply endorsement): Cursor IDE (Anysphere), Claude 4.5 Opus (Anthropic), GPT-5.1/GPT-4o (OpenAI), Gemini 3.0 Pro (Google DeepMind), Grok (xAI), Windsurf (Codeium), and Claude Code (Anthropic).

% ============================================================
% AUTHOR CONTRIBUTIONS
% ============================================================
\section*{Author Contributions}

Brandon Dent, MD conceived the project, designed the architecture, implemented the safety stack, conducted the evaluation, and wrote the manuscript.

% ============================================================
% REFERENCES
% ============================================================
\bibliographystyle{plain}
\bibliography{bibliography}

% ============================================================
% APPENDICES
% ============================================================
\appendix

\input{appendices/appendix_reproducibility}

\input{appendices/appendix_failure_taxonomy}

\input{appendices/appendix_safety_consort}

\end{document}
