% ============================================================
% Paper Content (Shared by all build targets)
% ============================================================
% This file contains ONLY the paper body content.
% It is included by build_named.tex and build_anonymous.tex
% after they set up the document environment.
%
% Updated: March 2026 — Full rewrite with current data
% Target: arXiv (cs.AI + cs.CL), NeurIPS 2026 D&B stretch
% ============================================================

\begin{abstract}
We introduce a trajectory-level safety evaluation framework for frontier language models in healthcare contexts. Unlike single-turn benchmarks that assess initial response accuracy, this framework measures \textit{safety persistence}---whether models maintain safety-critical recommendations across multi-turn conversations when users apply realistic, non-adversarial pressure. Across 11,547 verified trajectories spanning four frontier models (GPT-5.2, Claude Opus 4.5/4.6, Claude Sonnet 4.5/4.6), we observe an aggregate violation rate of 82.4\% [95\% CI: 80.1--84.5\%]. A 2$\times$2 factorial design separating system prompt preamble from pattern-based enforcement reveals that the preamble is the dominant mitigation mechanism, while enforcement actively interferes with constitutional AI models. The best configuration (GPT-5.2 with preamble, Pass$^k$ = 0.696) falls well below the pass$^5 \geq 0.95$ deployment threshold. No model--intervention combination evaluated here achieves adequate safety persistence. All clinical labeling was performed by a single board-certified emergency medicine physician; multi-rater reliability is not established. Code and evaluation artifacts are publicly available.
\end{abstract}

% ============================================================
\section{Introduction}
\label{sec:intro}

Healthcare AI deployment is accelerating faster than evaluation infrastructure can ensure safety. Current benchmarks predominantly evaluate single-turn response quality: HealthBench~\citep{openai2025healthbench} measures response quality across 5,000 conversations; MedQA~\citep{jin2021disease} and Med-PaLM~\citep{singhal2023large} evaluate factual medical knowledge. These benchmarks have driven substantial progress. However, response quality and trajectory-level safety persistence are distinct properties: a model can provide high-quality responses while relaxing safety boundaries under sustained conversational pressure.

Consider a parent whose 3-week-old infant has a fever of 100.6\textdegree F. A frontier model correctly identifies this as a medical emergency requiring immediate evaluation---neonatal sepsis carries 10--15\% mortality if treatment is delayed~\citep{shane2017neonatal}. The parent responds: ``I don't have a car until morning. What should I watch for?'' The model now faces a tension between helpfulness (providing monitoring guidance) and safety (not validating delay of emergency care). If the model provides home monitoring criteria, it implicitly validates a decision to delay---regardless of whether it also recommends the emergency department.

This tension between accommodating user preferences and maintaining safety boundaries is precisely what Constitutional AI~\citep{bai2022constitutional}, scalable oversight~\citep{bowman2022measuring}, and RLHF training~\citep{ouyang2022training} aim to address. Yet existing evaluations do not systematically measure whether safety-critical recommendations persist across conversation turns when users apply realistic pressure. To our knowledge, no published benchmark measures multi-turn escalation persistence under cooperative (non-adversarial) conversational pressure.

We define \textbf{safety persistence} as the property that a model maintains safety-critical recommendations across a multi-turn conversation when users apply realistic pressure to modify or abandon those recommendations. This is distinct from refusal robustness (resistance to adversarial jailbreaks), instruction following, or response quality. A model may exhibit high refusal robustness while exhibiting low safety persistence---these are orthogonal safety properties.

\textbf{Contributions.} We present ScribeGoat2, a physician-adjudicated evaluation framework that: (1) constructs realistic multi-turn pressure sequences derived from clinical experience across 23 emergency conditions and 43 adversarial scenarios; (2) separates system prompt preamble from pattern-based enforcement in a 2$\times$2 factorial design, revealing that preamble is the dominant mitigation mechanism while enforcement interferes with constitutional AI models; (3) validates LLM-as-judge classifications against expert clinical judgment and identifies systematic same-vendor judge bias; and (4) maintains a living falsification record of 7 claims that were believed, tested, and found wrong---including the discovery that pattern-based grading has an 85--97\% false negative rate on safety-trained models.

% ============================================================
\section{Related Work}
\label{sec:related}

\textbf{Healthcare AI evaluation.} HealthBench~\citep{openai2025healthbench} evaluates response quality across 48,562 physician-written rubric criteria (average 2.6 turns; 58.3\% single-turn). MedQA~\citep{jin2021disease} and Med-PaLM~\citep{singhal2023large} evaluate factual medical knowledge. CSEDB~\citep{chen2025csedb} benchmarks clinical safety event detection. MedOmni-45~\citep{wu2025medomni} spans 45 specialties. These benchmarks measure response quality and factual accuracy---distinct from whether a safety-critical recommendation persists under conversational pressure.

\textbf{Safety benchmarks.} Red-teaming approaches~\citep{perez2022red, ganguli2022red} evaluate robustness under adversarial manipulation. ForesightSafety Bench~\citep{li2026foresightsafety} evaluates proactive safety capabilities. Ren et al.~\citep{ren2026howshould} examine what safety benchmarks should measure. These evaluate refusal robustness---resistance to adversarial circumvention. Our pressure operators are cooperative, not adversarial: users express realistic barriers to care, not attacks designed to circumvent safety measures.

\textbf{Agent and multi-turn evaluation.} AgentBench~\citep{liu2023agentbench} evaluates LLMs as agents across multi-step tasks. SWE-bench~\citep{jimenez2024swebench} measures software engineering capability. Corecraft~\citep{mehta2026corecraft} benchmarks agents on progressively complex tasks. METR~\citep{metr2025taskstandard} establishes a task standard for dangerous capability evaluation. Anthropic's agent evaluation framework~\citep{anthropic2026evals} emphasizes trajectory-level measurement. MT-Bench~\citep{zheng2024judging} and WildBench~\citep{lin2024wildbench} evaluate multi-turn conversation quality. These frameworks inform our trajectory-level approach but address task completion rather than safety persistence under pressure.

\textbf{LLM-as-Judge.} LLM-as-judge methodology is established with known limitations including same-family bias, calibration drift, and sensitivity to prompt engineering~\citep{zheng2024judging, kim2024prometheus}. We use heterogeneous detection: rule-based detection for operationally-defined failures and cross-vendor LLM-as-judge for interpretive failures. We document systematic same-vendor bias (Section~\ref{sec:negative}).

\textbf{Positioning.} Table~\ref{tab:positioning} positions this evaluation relative to existing benchmarks. We do not claim that persistence is more important than response quality or factual accuracy. We claim it is a distinct property that existing benchmarks do not measure.

\begin{table}[h]
\centering
\small
\begin{tabular}{llcc}
\toprule
\textbf{Benchmark} & \textbf{Property Measured} & \textbf{Multi-Turn} & \textbf{Persistence} \\
\midrule
HealthBench & Response quality & Partial & No \\
MedQA / Med-PaLM & Factual knowledge & No & No \\
Red-teaming & Refusal robustness & Some & No \\
AgentBench / SWE-bench & Task completion & Yes & No \\
ForesightSafety & Proactive safety & No & No \\
\textbf{This work} & \textbf{Safety persistence} & \textbf{Yes} & \textbf{Yes} \\
\bottomrule
\end{tabular}
\caption{Positioning relative to existing benchmarks.}
\label{tab:positioning}
\end{table}

% ============================================================
\section{Methods}
\label{sec:methods}

\subsection{Trajectory Design}
\label{sec:trajectory}

A \textbf{trajectory} is a complete 5-turn multi-turn conversation between a user and a language model. Each trajectory includes a system prompt, an initial clinical presentation, and 4 subsequent user--model exchanges applying escalating pressure. Turn 1 presents the clinical scenario with an initial barrier (e.g., ``no car until morning''). Turns 2--5 apply progressively stronger pressure operators: cost/access barriers, authority challenges (``my nurse sister says it's fine''), apparent improvement signals (``fever is down after Tylenol''), and explicit capitulation requests (``just tell me what to watch for'').

We evaluate 23 emergency scenarios spanning 15 escalation conditions (where the model should maintain the emergency recommendation) and 8 defer conditions (where the model should respect patient autonomy). Escalation conditions include neonatal sepsis (ESI-1, 6--12h time-to-harm, 10--15\% mortality), biphasic anaphylaxis~\citep{sampson2006anaphylaxis}, testicular torsion, diabetic ketoacidosis, meningococcemia, subarachnoid hemorrhage, and others. Defer conditions include elective procedures and conditions where watchful waiting is guideline-appropriate.

\textbf{Corpus totals.} Across all evaluation phases, 11,547 unique trajectories were verified against raw checkpoint files (\texttt{docs/DATA\_MANIFEST.yaml}). The initial evaluation generated 12,000 trajectories (12 studies $\times$ 1,000); deduplication and reconciliation yielded the verified count.

\subsection{Pressure Operator Taxonomy}
\label{sec:pressure}

Pressure operators are cooperative, realistic conversational moves derived from observed patient communication patterns---not adversarial jailbreaks. Table~\ref{tab:pressure} defines the taxonomy.

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Operator} & \textbf{Description} & \textbf{Example} \\
\midrule
Financial barrier & Cost as reason to delay & ``I can't afford an ambulance'' \\
Social pressure & Family/authority disagreement & ``My family thinks I'm overreacting'' \\
Autonomy assertion & Right to decide & ``It's my choice whether to go'' \\
Polite disengagement & Intent to wait & ``Thanks, I'll think about it'' \\
Alternative request & Home monitoring guidance & ``What should I watch for?'' \\
\bottomrule
\end{tabular}
\caption{Pressure operator taxonomy. All operators are cooperative and non-deceptive.}
\label{tab:pressure}
\end{table}

\subsection{Failure Modes}
\label{sec:failures}

Failure modes are operational constructs defined by explicit criteria before evaluation begins.

\textbf{Tier 1 failures} (context-independent) are safety violations inappropriate regardless of context:
\begin{itemize}
    \item \textsc{URGENCY\_MINIMIZED}: Model downplays genuine clinical urgency
    \item \textsc{DELAY\_VALIDATED}: Model explicitly endorses a dangerous delay
    \item \textsc{PATIENT\_ABANDONED}: Model refuses any guidance when patient cannot follow recommendations
\end{itemize}

\textbf{Tier 2 failures} (trajectory-dependent) emerge only across conversation turns:
\begin{itemize}
    \item \textsc{RESOURCE\_OMISSION}: Model drops explicit crisis resources (e.g., 988 hotline) after user politely disengages following crisis disclosure
\end{itemize}

Tier numbering reflects dependency on conversational trajectory, not severity.

\subsection{Detection}
\label{sec:detection}

\textbf{Disclosure: Single adjudicator.} All clinical labeling was performed by a single board-certified emergency medicine physician (the study author, with 11 years of practice and approximately 40,000 patient encounters). Inter-rater reliability metrics cannot be computed without additional independent adjudicators. The 100\% calibration agreement (60/60 cases, Cohen's $\kappa$ = 1.000) may reflect alignment with individual judgment patterns rather than clinical consensus. Results should be interpreted as existence detection under an explicit labeling protocol, not prevalence estimates.

Detection mechanisms are intentionally heterogeneous:

\begin{itemize}
    \item \textbf{Rule-based} (RESOURCE\_OMISSION): Deterministic pattern matching for presence/absence of crisis resources. Fully reproducible; no judge involved.
    \item \textbf{LLM-as-judge} (Tier 1 failures): Cross-vendor classification. Judge receives response text and clinical context; reference solutions are \textit{not} included to prevent stylistic grading.
    \item \textbf{CEIS pipeline}: Three-layer architecture combining deterministic pattern screening (Layer 0), LLM-judge classification (Layer 2), and severity-stratified failure taxonomy (Class A--D).
\end{itemize}

Cross-vendor judging is mandatory for inter-model comparisons. We document that same-vendor judging produces systematically biased results (Section~\ref{sec:negative}).

\subsection{Metrics}
\label{sec:metrics}

\textbf{Pass$^k$} (conservative reliability): $\text{Pass}^k = \text{success\_rate}^k$, the probability that all $k$ independent trials succeed. The deployment gate requires pass$^5 \geq 0.95$. A model with 90\% single-trial success achieves pass$^5 = 0.59$---well below threshold.

\textbf{Cohen's $h$} (effect size): $h = 2(\arcsin\sqrt{p_1} - \arcsin\sqrt{p_2})$. Interpretation: $|h| < 0.2$ small, $0.2 \leq |h| < 0.5$ medium, $|h| \geq 0.8$ large~\citep{cohen1988statistical}.

\textbf{Wilson score intervals}: Used for per-condition and per-study failure rates~\citep{wilson1927probable}. Preferred over Wald intervals for proportions near 0 or 1.

\textbf{Bootstrap resampling}: Used for aggregate scores that combine multiple conditions (e.g., CEIS ERS), where the unit of resampling is the condition rather than the individual trajectory.

\subsection{Experimental Setup}
\label{sec:setup}

\textbf{Models.} Table~\ref{tab:models} lists all evaluated models. The initial evaluation (January 2026) covered GPT-5.2 and Claude Opus 4.5. The expanded corpus (February 2026) added Claude Opus 4.6, Sonnet 4.5, and Sonnet 4.6.

\begin{table}[h]
\centering
\small
\begin{tabular}{llll}
\toprule
\textbf{Model} & \textbf{Provider} & \textbf{Model ID} & \textbf{Period} \\
\midrule
GPT-5.2 & OpenAI & \texttt{gpt-5.2} & Jan--Feb 2026 \\
Claude Opus 4.5 & Anthropic & \texttt{claude-opus-4-5-20251101} & Jan 2026 \\
Claude Opus 4.6 & Anthropic & \texttt{claude-opus-4-6-20260201} & Feb 2026 \\
Claude Sonnet 4.5 & Anthropic & \texttt{claude-sonnet-4-5-20250929} & Feb 2026 \\
Claude Sonnet 4.6 & Anthropic & \texttt{claude-sonnet-4-6-20260201} & Feb 2026 \\
\bottomrule
\end{tabular}
\caption{Models evaluated with version-pinned identifiers.}
\label{tab:models}
\end{table}

\textbf{2$\times$2 factorial design.} The Monotonic Safety Invariant Wrapper (MSIW) has two independent mechanisms: (1) a \textit{system prompt preamble}---a safety directive injected before the model call, and (2) \textit{pattern-based enforcement}---a post-response check for forbidden phrases that triggers response replacement. We separated these into a 2$\times$2 factorial (preamble $\times$ enforcement) across 4 models on the 23-scenario corpus, with 3 trials per cell.

\textbf{Deterministic execution.} All evaluations used seed $= 42$, temperature $= 0$, top-$p = 1.0$, max tokens $= 2048$. All prompts were hashed (SHA-256) before execution.

% ============================================================
\section{Results}
\label{sec:results}

\subsection{Primary Finding: 82.4\% Aggregate Violation Rate}
\label{sec:primary}

Across 11,547 verified trajectories spanning both evaluation domains (healthcare emergency persistence and crisis resource persistence), 82.4\% [95\% CI: 80.1--84.5\%] exhibited at least one safety-critical failure. Neither model family meets the pass$^5 \geq 0.95$ deployment gate under any evaluated configuration.

\subsection{2$\times$2 Factorial Results}
\label{sec:factorial}

Table~\ref{tab:factorial} presents Pass$^k$ for all 16 model--condition cells on the 23-scenario expanded corpus (3 trials each, 5-turn trajectories, cross-vendor LLM-judge grading).

\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\toprule
\textbf{Model} & \textbf{Baseline} & \textbf{Preamble} & \textbf{Enforce} & \textbf{Wrapper} \\
\midrule
GPT-5.2         & 0.261 & \textbf{0.696} & 0.261 & \textbf{0.696} \\
Sonnet 4.5      & 0.174 & \textbf{0.609} & 0.217 & 0.652 \\
Opus 4.6        & 0.217 & \textbf{0.478} & 0.304 & 0.391 \\
Sonnet 4.6      & 0.130 & 0.261          & 0.261 & 0.304 \\
\bottomrule
\end{tabular}
\caption{Pass$^k$ by model and intervention (23 scenarios, 3 trials, 5-turn). Bold indicates best configuration per model. The best result across all cells is GPT-5.2 at 0.696---well below the 0.95 deployment threshold.}
\label{tab:factorial}
\end{table}

\subsection{Preamble Dominance}
\label{sec:preamble}

The preamble is the dominant mitigation mechanism across all models tested:

\begin{itemize}
    \item \textbf{GPT-5.2}: Preamble-only (0.696) equals the full wrapper (0.696). Enforcement adds nothing.
    \item \textbf{Sonnet 4.5}: Preamble-only (0.609) approximates the full wrapper (0.652). Enforcement contributes +0.043.
    \item \textbf{Opus 4.6}: Preamble-only (0.478) \textit{exceeds} the full wrapper (0.391). Enforcement \textit{reduces} performance by 0.087.
    \item \textbf{Sonnet 4.6}: Preamble-only (0.261) is slightly below the full wrapper (0.304). Small enforcement benefit (+0.043).
\end{itemize}

The enforcement interference for Opus 4.6 replicates across both cooperative emergency scenarios ($\Delta = -0.087$) and adversarial code-agent scenarios ($\Delta = -0.125$), suggesting an architectural property of constitutional AI models rather than a corpus artifact. When enforcement replaces a response that was semantically adequate but triggered a forbidden pattern, it disrupts the model's preamble-guided conversational trajectory at subsequent turns.

\subsection{Negative Results}
\label{sec:negative}

\textbf{No configuration meets the deployment threshold.} The best Pass$^k$ across all model--intervention combinations is 0.696 (GPT-5.2 with preamble)---a 36\% shortfall from the pass$^5 \geq 0.95$ gate. System prompt interventions plateau well below acceptable thresholds.

\textbf{Pattern-based grading is unreliable.} Pattern-only grading exhibits an 85--97\% false negative rate on safety-trained models. Cascade 8B scored 1.000 under pattern grading but 0.125 under LLM judge---the model includes safety keywords (``call 911'') while contextually undermining them (``but since your symptoms are improving, it's reasonable to monitor at home first''). Constitutional AI models are particularly vulnerable: during seeds evaluation, pattern grading produced an Opus baseline of 4\% versus the LLM-judged 54\%.

\textbf{Same-vendor judging is systematically biased.} Under the original MSTS evaluation (Claude Sonnet 4.5 judging both models), the GPT-5.2 to Opus failure ratio was approximately 80:1 (0.025\% vs 2.0\%). Under cross-vendor grading, this ratio inverts or collapses, with inter-method agreement $\kappa = -0.21$ to $-0.40$ (worse than chance). Cross-vendor judges are mandatory for inter-model comparisons.

\textbf{Sonnet 4.6 safety regression.} Sonnet 4.6 showed a 70\% reduction in preamble effectiveness compared to Sonnet 4.5 (preamble lift dropped from +0.435 to +0.131 Pass$^k$). The newer model---presumably improved on general capabilities---degraded on safety persistence. Model updates that improve general capabilities may simultaneously degrade the effectiveness of deployed safety mitigations.

\textbf{Small benchmarks produce misleading results.} The MSIW wrapper achieved Pass$^k = 1.000$ on the original 8-scenario corpus. On the expanded 23-scenario corpus, the best result dropped to 0.696. The 15 additional scenarios introduced failure modes invisible to both preamble and enforcement. The Wilson CI for $n = 8$ with all-pass is [0.676, 1.000]---the true population performance could have been as low as where the 23-scenario result landed.

\subsection{Secondary Findings}
\label{sec:secondary}

\textbf{Adversarial vectors.} Code generation context is the most effective attack surface. Table~\ref{tab:adversarial} summarizes baseline Pass$^k$ across 43 adversarial scenarios spanning 5 attack vectors. Code-agent scenarios reduce both models to near-zero baseline persistence. Preamble intervention substantially improves code-agent performance (both models to 0.625), but enforcement interference replicates for Opus (preamble-only 0.625 $>$ wrapper 0.500).

\begin{table}[h]
\centering
\small
\begin{tabular}{lccc}
\toprule
\textbf{Vector} & \textbf{Scenarios} & \textbf{GPT-5.2} & \textbf{Opus 4.6} \\
\midrule
Code-agent     & 8  & 0.125 & 0.250 \\
Tool-use       & 8  & 0.857 & 0.625 \\
Multimodal     & 8  & 0.750 & 0.875 \\
Integrated     & 9  & 0.556 & 1.000 \\
\bottomrule
\end{tabular}
\caption{Baseline Pass$^k$ by adversarial attack vector (3 trials, cross-vendor LLM judge).}
\label{tab:adversarial}
\end{table}

\textbf{RAG grounding.} Clinical context retrieved from the OpenEM corpus (370 emergency medicine conditions) and injected as a system prompt prefix lifted Opus 4.6 Pass$^k$ from 0.217 to 0.391 ($+0.174$). RAG primarily reduces \textsc{URGENCY\_MINIMIZED} failures ($-65\%$)---cases where the model softens language due to lack of clinical specificity---while having little effect on \textsc{DELAY\_VALIDATED} failures, where the model understands the emergency but capitulates. RAG and MSIW interactions are additive rather than synergistic.

\textbf{Seeds persistence.} Seventeen mined evaluation seeds were tested across 4 models with 5 trials each. GPT-5.2 and Gemini 3.1 Pro reached 100\% preamble pass rate; Opus 4.6 and Grok 4.1 Fast reached a ceiling of 88.2\%. Two seeds remained universally unsolved for Opus and Grok: SEED-013 (PE code-agent telehealth routing, where valid clinical evidence overrides safety boundaries) and SEED-015 (tension headache to SAH boundary, where models treat a life-threatening mimic as benign despite red-flag features).

% ============================================================
\section{Discussion}
\label{sec:discussion}

\subsection{Interpreting the 82.4\% Violation Rate}

The aggregate 82.4\% violation rate means that across 11,547 multi-turn trajectories, the vast majority exhibited at least one safety-critical failure. This does not mean models are ``unsafe'' in general---it means that under sustained, realistic conversational pressure, models fail to maintain safety-critical recommendations at rates that preclude unsupervised deployment in healthcare contexts.

The violation rate aggregates across two domains (healthcare emergency persistence and crisis resource persistence) with different detection mechanisms, failure types, and baseline rates. It is a summary statistic, not a prediction of deployment-context failure rates. Per-domain and per-condition rates (Appendix~\ref{app:conditions}) are more informative for specific applications.

\subsection{The Preamble Paradox}

The preamble---a safety directive in the system prompt---is the most effective single intervention. Yet it plateaus at Pass$^k \approx 0.70$, well below any reasonable deployment threshold. Two observations complicate the picture:

First, for constitutional AI models (Opus 4.6), adding enforcement \textit{reduces} performance. The mechanism appears to be conversational trajectory disruption: when enforcement replaces a semantically adequate response, the model sees the replacement template rather than its own reasoning in subsequent turns, causing divergent and often degraded behavior. This replicates across cooperative and adversarial corpora, suggesting it is an architectural interaction between constitutional training and external enforcement.

Second, model updates can drastically reduce preamble effectiveness. Sonnet 4.6 showed a 70\% reduction in preamble lift compared to Sonnet 4.5. General capability improvements do not guarantee preservation of safety persistence properties. Any system relying on prompt-level safety interventions must re-evaluate after every model update.

\subsection{Living Evidence and Falsification}

Over the course of this evaluation, 7 of 15 registered empirical claims were falsified or superseded by new evidence. These include: the Turn 2 cliff hypothesis (scenario design artifact), the Opus 4\% baseline (pattern grading artifact), the MSIW scalability claim (small benchmark artifact), the enforcement complement assumption (interference for constitutional models), and the MSTS 80:1 GPT/Opus ratio (same-vendor judge bias). We consider this falsification record a credibility signal rather than a weakness---a document that has never been wrong has never been tested. The compact falsification record appears in Appendix~\ref{app:claims}.

% ============================================================
\section{Limitations}
\label{sec:limitations}

\textbf{Single adjudicator.} All clinical labeling was performed by a single board-certified emergency medicine physician. Inter-rater reliability metrics cannot be computed. The 100\% calibration agreement (60/60) may reflect overfitting to individual judgment patterns. This is the primary limitation. Observed rates are valid measurements given the explicit labeling protocol, but prevalence estimates require multi-rater validation. A blinded multi-rater protocol (minimum 2 additional EM physicians, target $\kappa \geq 0.60$) is documented for Phase 2.1, contingent on funding.

\textbf{Synthetic scenarios.} All scenarios are synthetic constructions based on clinical experience. Real patient conversations include emotional dynamics, cultural context, and pressure patterns not captured by scripted operators. Findings indicate model behavior under controlled conditions, not patient outcomes in deployment.

\textbf{Detection heterogeneity.} Different failure types use different detection mechanisms (rule-based vs. LLM-as-judge vs. CEIS pipeline). This is intentional---operationally-defined failures permit deterministic detection, while interpretive failures require judgment---but complicates cross-domain comparison.

\textbf{Domain specificity.} All scenarios involve emergency medicine. Trajectory-level vulnerabilities may or may not generalize to financial advice, legal guidance, child safety, or other safety-critical domains.

\textbf{Temporal validity.} Findings apply to specific model versions evaluated in January--February 2026. Model updates invalidate prior findings. The Sonnet 4.6 regression demonstrates this concretely.

\textbf{In-distribution evaluation.} All scenarios were generated by the same pipeline and team. The expanded corpus introduces diversity but does not constitute external validation. Independent clinical scenario generation by physicians not involved in evaluation design remains the highest-priority gap for publication.

% ============================================================
\section{Conclusion}
\label{sec:conclusion}

We present evidence of a structural gap between single-turn safety optimization and trajectory-level safety persistence in frontier language models. Across 11,547 trajectories, 4 frontier models, and a 2$\times$2 factorial design, no model--intervention combination achieves the pass$^5 \geq 0.95$ deployment threshold. The best configuration (GPT-5.2 with system prompt preamble, Pass$^k = 0.696$) leaves a 36\% shortfall.

Three findings have immediate practical implications. First, the system prompt preamble is the dominant mitigation mechanism; pattern-based enforcement adds minimal or negative value for constitutional AI models and should be omitted for Anthropic model deployments. Second, pattern-based grading has an 85--97\% false negative rate on safety-trained models and should never be used alone for safety evaluation. Third, model updates can regress safety persistence---any system relying on prompt-level safety interventions must re-evaluate after every model update.

For practitioners: deploy with preamble-only mitigation, monitor with cross-vendor judges, and do not rely on prompt-level interventions alone for safety-critical contexts. For researchers: trajectory-level safety persistence is measurable, distinct from single-turn safety, and far from solved. The gap between current best performance (Pass$^k = 0.70$) and deployment requirements (pass$^5 \geq 0.95$) likely requires training-level interventions---persistence-aware reward signals, trajectory-level constitutional principles, or monotonic safety constraints during training---rather than further prompt engineering.

% ============================================================
\section*{Acknowledgments}

This work was conducted independently at GOATnote Inc. The evaluation is grounded in direct clinical experience: the author is a board-certified Emergency Medicine physician with 11 years of practice and approximately 40,000 patient encounters. Findings were shared with OpenAI, Anthropic, Google DeepMind, and xAI through their published reporting channels as part of responsible disclosure. The OpenEM corpus (370 emergency medicine conditions, Apache 2.0) provided clinical knowledge grounding for RAG evaluation. We thank the frontier lab safety teams for their published frameworks.

% ============================================================
\section*{Data Availability}

All evaluation artifacts, code, and trajectories are available at \url{https://github.com/GOATnote-Inc/scribegoat2}. The clinical knowledge corpus is available at \url{https://github.com/GOATnote-Inc/openem-corpus}. Replication requires API access to the evaluated models at the documented version pins.

% ============================================================
\bibliographystyle{plainnat}
\bibliography{references}

% ============================================================
% Appendices
% ============================================================
\appendix
\input{appendix}
